Advanced Normalization Tools (ANTS)

### 2009-06-11 Documentation/Tex/ants.tex
add metric.tex as a separate tex file, may be removed later
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@22 fd2da0a4-d968-48df-9cb5-d8c496287cbf
need Figures/r16...
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@24 fd2da0a4-d968-48df-9cb5-d8c496287cbf

\subsection{The ANTS Executable}
The ANTS program itself is the central program encapsulating
normalization/registration functionality.  Its main output is an
affine transform file and a deformation field, potentially with
inverse.  Options to ANTS allows the user to navigate the similarity
and transformation options that are available.  ANTS takes multiple
similarity and optimization criteria as options.  The program is
wrapped in \verb ants.sh for normalization with ``out of the box''
parameters and in \verb buildtemplateparallel.sh for 
computationally distributed optimal template construction.

@@

\subsection{The ImageMath Executable} This is a multi-purpose program that has the following syntax:  ``ImageMath ImageDimension outputfilename   Operation  InputFileName  parameters ``.  Most basic scalar image operations -- and some tensor operations -- may be performed with this program.  Some operations output text files, some output images and some output only to the terminal.   ImageMath allows one to multiply images together (m), to negate images (Neg), to take an image to a power (pow), to test the invertibility of transformations (InvId), to compute the fractional anisotropy of an image (TensorFA) and to perform Markov Random Field segmentation with our without priors (Segment).   Many other operations are available.  Like all other ANTS programs, one may call ImageMath from the command line to see all of its options.   ImageMath is used heavily in ANTS scripts.

@@

\noindent{\bf Elastic/Vector Space Transformations.} If we assume no affine transformation, then 
an elastic transformation involves computing a mapping from image $I(\x)$ to image $J(\x)$ 
through a deformation field $\disp( \x )$.  
\begin{wrapfigure}{r}{0.5\textwidth}
\label{fig:pull}
\includegraphics[width=0.45\textwidth]{./Figures/PullBack.pdf} 
\end{wrapfigure}  
The deformation is defined in the physical space of the image 
and dictates the positional difference between corresponding features in the two images.  
Thus, if a feature defined at $I(\x)$ matches a feature in $J$ at position $\y$ then the deformation 
field at $\x$ should give $\disp ( \x )=\y-\x$.   
Such a deformation field may be applied to deform 
image $J$ into image $I$ by composing the mapping $J_{deformed}(\x)=J(\x+\disp(\x))$.  
In a perfect world, then $I(\x)=J_{deformed}(\x)$, though this is rarely the case.  Figure~\ref{fig:pull} 
visualizes the deformation of an image under this standard model. 
Gradient descent optimization of an elastic mapping may be summarized (crudely) as: 
\begin{align}
\text{Compute the similarity gradient:   }  \nabla E = \partial_{\disp} \Pi(I,J(\x+\disp(\x))). \nonumber \\
\text{Update the deformation field:  }  \disp(\x) \leftarrow \disp(\x)+\delta \nabla E. \nonumber \\ 
\text{Regularize the deformation field:  }  \disp(\x) \leftarrow G_\sigma \star \disp(\x),
\end{align}
where $\Pi$ is the similarity, $\delta$ is a gradient step length and $G_\sigma$ is a gaussian smoother.
This optimization is captured in the following ANTS command:
\begin{verbatim}
ANTS 3 -m PR[AtlasHead.nii,crAnatomical.nii,1,4] -r Gauss[0,3]  
-o ElasticTest.nii.gz -i 30x20x10  -t Elast[1.5] 
\end{verbatim}
Here, we use the PR metric (a cross-correlation implementation) with
window radius 4, weight 1 and gradient step length 1.5.  The
optimization will be performed over three resolutions with a maximum
of 30 iterations at the coarsest level, 20 at the next coarsest and 10
at the full resolution.  We use a Gaussian regularizer with a sigma of
3 that operates only on the deformation field and not on the similarity gradient, as 
0 is passed as the first parameter to \verb Gauss .  One may see the
correspondence, yet again, between the ANTS call and the optimization
scheme.  The optimization will stop when either the energy cannot be 
smoothly minimized or the maximum number of iterations is reached. 
BSpline regularization is also available in ANTS -- see the DMFFD 
section below.

@@

>>>>

\subsection{The ANTS Executable}
The ANTS program itself is the central program encapsulating
normalization/registration functionality.  Its main output is an
affine transform file and a deformation field, potentially with
inverse.  Options to ANTS allows the user to navigate the similarity
and transformation options that are available.  ANTS takes multiple
similarity and optimization criteria as options.  The program is
wrapped in \verb ants.sh ~for normalization with ``out of the box''
parameters and in \verb buildtemplateparallel.sh ~for 
computationally distributed optimal template construction.

@@

\subsection{The ImageMath Executable} This is a multi-purpose program that has the following syntax:  ``ImageMath ImageDimension outputfilename   Operation  InputFileName  parameters ``.  Most basic scalar image operations -- and some tensor operations -- may be performed with this program.  Some operations output text files, some output images and some output only to the terminal.   ImageMath allows one to multiply images together (m), to negate images (Neg), to take an image to a power (pow), to test the invertibility of transformations (InvId), to compute the fractional anisotropy of an image (TensorFA) and to perform Markov Random Field segmentation with or without priors (Segment).   Many other operations are available.  Like all other ANTS programs, one may call ImageMath from the command line to see all of its options.   ImageMath is used heavily in ANTS scripts.

@@

\noindent{\bf Elastic/Vector Space Transformations.} If we assume no affine transformation, then 
an elastic transformation involves computing a mapping from image $I(\x)$ to image $J(\x)$ 
through a deformation field $\disp( \x )$.  
\begin{figure}
\label{fig:pull}
\center \includegraphics[width=0.75\textwidth]{./Figures/PullBack.pdf} 
\itkcaption[ANTS Deformation]{A deformation in a digital domain.}
\end{figure}  
The deformation is defined in the physical space of the image 
and dictates the positional difference between corresponding features in the two images.  
Thus, if a feature defined at $I(\x)$ matches a feature in $J$ at position $\y$ then the deformation 
field at $\x$ should give $\disp ( \x )=\y-\x$.   
Such a deformation field may be applied to deform 
image $J$ into image $I$ by composing the mapping $J_{deformed}(\x)=J(\x+\disp(\x))$.  
In a perfect world, then $I(\x)=J_{deformed}(\x)$, though this is rarely the case.  Figure~\ref{fig:pull} 
visualizes the deformation of an image under this standard model. 
Gradient descent optimization of an elastic mapping may be summarized (crudely) as: 
\begin{align}
\text{Compute the similarity gradient:   }  \nabla E = \partial_{\disp} \Pi(I,J(\x+\disp(\x))). \nonumber \\
\text{Update the deformation field:  }  \disp(\x) \leftarrow \disp(\x)+\delta \nabla E. \nonumber \\ 
\text{Regularize the deformation field:  }  \disp(\x) \leftarrow G_\sigma \star \disp(\x),
\end{align}
where $\Pi$ is the similarity, $\delta$ is a gradient step length and $G_\sigma$ is a gaussian smoother.
This optimization is captured in the following ANTS command:
\begin{verbatim}
ANTS 3 -m PR[AtlasHead.nii,crAnatomical.nii,1,4] -r Gauss[0,3]  
-o ElasticTest.nii.gz -i 30x20x10  -t Elast[1.5] 
\end{verbatim}
Here, we use the PR metric (a cross-correlation implementation) with
window radius 4, weight 1 and gradient step length 1.5.  The
optimization will be performed over three resolutions with a maximum
of 30 iterations at the coarsest level, 20 at the next coarsest and 10
at the full resolution.  We use a Gaussian regularizer with a sigma of
3 that operates only on the deformation field and not on the similarity gradient, as 
0 is passed as the first parameter to \verb Gauss .  One may see the
correspondence, yet again, between the ANTS call and the optimization
scheme.  The optimization will stop when either the energy cannot be 
smoothly minimized or the maximum number of iterations is reached. 
BSpline regularization is also available in ANTS -- see the DMFFD 
section below.

@@

\subsection{Choosing a Metric}
\input{metric.tex}
